---
title: "Parent material and climate interact to control soil carbon dynamics on timescales from years to centuries"
author: J. Beem-Miller
date: \textit{\today}
output:
  pdf_document:
    latex_engine: xelatex
    toc: yes
    toc_depth: 3
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
  html_notebook:
    css: "custom.css"
    toc: yes
    toc_depth: 2
header_includes:
- \usepackage[utf8]{inputenc}
- \usepackage{float}
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.align = 'center', dev = c('cairo_pdf', 'png'), fig.width = 6.5,
                      fig.height = 3.5)
options(scipen = 5)
# load page breaks fx
source("./utilities/page_break_Rmd.R")
```

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(SoilR)
library(openxlsx)
library(ISRaD)
library(lme4)
library(lmerTest)
library(emmeans)
library(gt)
library(scales)
```

```{r plot-funs}
# color palettes for ECO & PM
warm <- "#BF812D"
cool <- "#80CDC1"
cold <- "#01665E"
granite <- "#9daba9"
andesite <- "#382dbf"
basalt <- "#bf382d"

# emmimp plots
emmip.plot.fx <- function(lyr_bot, Type, var, interact_var = NULL) {
  
  # get data, set limits, and define ylab
  if (Type == "Bulk") {
    data <- sra.01.09.19.d.df[sra.01.09.19.d.df$lyr_bot == lyr_bot, ]
    yLab <- expression(Delta*''^14*'C (‰)')
    lim <- c(-130, 165)
    brks <- c(2001, 2009, 2019)
    shp <- 16
  } else {
    data <- sra.01.19.inc.d.df[sra.01.19.inc.d.df$lyr_bot == lyr_bot, ]
    yLab <- expression(Delta*''^14*'C-CO'[2]*' (‰)')
    lim <- c(-190, 170)
    brks <- c(2001, 2019)
    shp <- 1
  }
  
  # model
  mod <- lm(d14c ~ PM * Year * ECO, data)
  
  # set scales
  if (var == "PM") {
    nm <- "Parent material"
    vals <- c("andesite" = andesite, "basalt" = basalt, "granite" = granite)
  } else {
    nm <- "Climate"
    vals <- c("warm" = warm, "cool" = cool, "cold" = cold)
  }
  
  # define depth and var
  depth <- ifelse(lyr_bot == 10, "0-10cm", ifelse(lyr_bot == 20, "10-20cm", "20-30cm"))
  quo_var <- sym(var)
  quo_var_f <- ifelse(var == "PM", sym("ECO"), sym("PM"))
  
  # run emmip
  if (is.null(interact_var)) {
    if (var == "PM") {
      df <- emmip(mod, PM ~ Year, CIs = TRUE, at = list(Year = brks))$data 
    } else {
      df <- emmip(mod, ECO ~ Year, CIs = TRUE, at = list(Year = brks))$data
    }
  } else {
    if (var == "PM") {
      df <- emmip(mod, PM ~ Year | ECO, CIs = TRUE, at = list(Year = brks))$data
    } else {
      df <- emmip(mod, ECO ~ Year | PM, CIs = TRUE, at = list(Year = brks))$data
    }
  }
  ggplot(df, aes(xvar, yvar)) +
    geom_line(data = atm.14c, aes(Year, d14c, linetype = atm), color = "black") +
    geom_line(aes(color = !! quo_var), size = 1.5) +
    geom_ribbon(aes(ymin = LCL, ymax = UCL, fill = !! quo_var), alpha = .1) +
    geom_point(data = data, aes(Year, d14c, color = !! quo_var), 
               shape = shp, size = 2, alpha = .8) +
    scale_linetype_manual(name = NULL, 
                          values = c("atmosphere" = 3)) +
    scale_color_manual(name = nm,
                       values = vals) +
    scale_fill_manual(name = nm,
                      values = vals) +
    scale_x_continuous(limits = c(2000, 2020), breaks = brks) +
    coord_cartesian(ylim = lim) +
    ylab(yLab) +
    ggtitle(paste(Type, depth, sep = ", ")) +
    facet_grid(cols = vars(!! quo_var_f)) +
    theme_bw() +
    theme(panel.grid = element_blank(),
          plot.title = element_text(hjust = 0.5),
          axis.title.x = element_blank())
}
```

```{r atm-14c}
load("atm.14c.RData")
atm.14c <- atm.14c %>% rename(Year = year) %>% mutate(atm = "atmosphere")
```

```{r load-min}
load("ras18.sp.df.RData")
```

```{r bulk-data-wrangling}
# load data
load("sra.19.01.09.RData")
load("sra.19.01.rep.ls.RData")
load("sra.2019.ls.RData")

## change over time
# merge '01 and '19 data
sra.19.01.d <- merge(sra.19.01.09[sra.19.01.09$Year == 2001 & sra.19.01.09$lyr_bot, ],
                     sra.19.01.09[sra.19.01.09$Year == 2019 & sra.19.01.09$lyr_bot < 31, ],
                     by = c("PM", "ECO", "PMeco", "lyr_top", "lyr_bot"),
                     suffixes = c("_01", "_19")) 
sra.19.01.d$dd14c_01_19 <- sra.19.01.d$d14c_01 - sra.19.01.d$d14c_19
sra.19.01.d$sd14c_01_19 <- sqrt(sra.19.01.d$d14c_sd_01^2/3 + sra.19.01.d$d14c_sd_19^2/3)

## summarize change '01-'19
# PM
sra.19.01.dsum.PM <-
  sra.19.01.d %>%
  filter(lyr_bot < 31) %>%
  group_by(PM, lyr_bot) %>%
  summarize(across(dd14c_01_19, .fn = list(mean = mean, sd = sd), na.rm = TRUE))
# ECO
sra.19.01.dsum.ECO <-
  sra.19.01.d %>%
  filter(lyr_bot < 31) %>%
  group_by(ECO, lyr_bot) %>%
  summarize(across(dd14c_01_19, .fn = list(mean = mean, sd = sd), na.rm = TRUE))

## make long data frame of rep data
# 2001 = sra.19.01.rep.ls (splined to '19 depths)
sra.19.01.rep.df <- bind_rows(lapply(sra.19.01.rep.ls, function(PMeco) {
  df <- bind_rows(lapply(PMeco, function(r) {
    fm <- r[!is.na(r)]
    lyr_bot <- seq(10, length(fm) * 10, by = 10)
    return(data.frame(fm = fm, lyr_bot = lyr_bot))
    }), .id = "PMeco_rep")
  df$pro_rep <- substr(df$PMeco_rep, nchar(df$PMeco_rep), nchar(df$PMeco_rep))
  return(df[ , c("fm", "lyr_bot", "pro_rep")])
}), .id = "PMeco")
sra.19.01.rep.df$Year <- 2001
sra.19.01.rep.df$PM <- substr(sra.19.01.rep.df$PMeco, 1, 2)
sra.19.01.rep.df$ECO <- substr(sra.19.01.rep.df$PMeco, 3, 4)
# 2019 = sra.2019.ls
sra.2019.rep.df <- bind_rows(sra.2019.ls, .id = "PMeco")
sra.2019.rep.df <- sra.2019.rep.df[sra.2019.rep.df$lyr_bot < 31, ]

# 2009 (splined to '19 depths)
sra.09.19.df <- sra.19.01.09[sra.19.01.09$Year == "2009", 
                             c("PMeco", "fm", "lyr_bot", "Year", "PM", "ECO")] %>%
  mutate(pro_rep = 1) %>%
  filter(lyr_bot < 31)

# merge
sra.01.09.19.d.df <- rbind(
  sra.2019.rep.df[ , match(names(sra.19.01.rep.df), names(sra.2019.rep.df))],
  sra.09.19.df,
  sra.19.01.rep.df) %>%
  left_join(., ras18.sp.df[ , c("Al_ox", "Al_py", "Fe_dc", "Fe_ox", "lyr_bot", "PM", "ECO")], 
  by = c("lyr_bot", "PM", "ECO")) %>%
  mutate(PM = factor(recode(PM, `AN` = "andesite", `BS` = "basalt", `GR` = "granite")),
         ECO = factor(recode(ECO, `pp` = "warm", `wf` = "cool", `rf` = "cold"),
                      levels = c("warm", "cool", "cold")),
         year = as.numeric(Year) - 2000,
         Year = as.numeric(Year))
sra.01.09.19.d.df$d14c <- convert_fm_d14c(fm = sra.01.09.19.d.df$fm,
                                          obs_date_y = sra.01.09.19.d.df$Year, verbose = FALSE)

# by depth
sra.01.09.19.d.10df <- sra.01.09.19.d.df[sra.01.09.19.d.df$lyr_bot == 10, ]
sra.01.09.19.d.20df <- sra.01.09.19.d.df[sra.01.09.19.d.df$lyr_bot == 20, ]
sra.01.09.19.d.30df <- sra.01.09.19.d.df[sra.01.09.19.d.df$lyr_bot == 30, ]
```

```{r resp-data-wrangling}
# load data
load("sra.19.01.inc.df.RData")
load("sra.2019.inc.df.RData")
load("sra.2001.inc.df2.RData")
load("sra.2019.inc.df2.RData")

# rbind '01 and '19 data
sra.01.19.inc.d.df <- rbind(cbind(sra.2001.inc.df2, 
                                  rep = rep(c("a", "b"), each = 27),
                                  Year = 2001),
                            sra.2019.inc.df2[ , c("PM", "ECO", "rep", "lyr_bot", "Year", "d14c")]) %>%
  left_join(., ras18.sp.df[ , c("Al_ox", "Al_py", "Fe_dc", "Fe_ox", "lyr_bot", "PM", "ECO")], 
  by = c("lyr_bot", "PM", "ECO")) %>%
  mutate(PM = factor(recode(PM, `AN` = "andesite", `BS` = "basalt", `GR` = "granite")),
         ECO = factor(recode(ECO, `pp` = "warm", `wf` = "cool", `rf` = "cold"),
                      levels = c("warm", "cool", "cold")),
         year = Year - 2000)

## change over time
# merge '01 and '19 data
sd.inc.fx <- function(df) {
  apply(df[ , c("d14c_min", "d14c_max")], 1, sd)
}
sra.19.01.inc.d.df <- merge(
  cbind(sra.19.01.inc.df, sd = sd.inc.fx(sra.19.01.inc.df))[ , c("PM", "ECO", "PMeco", "lyr_bot", "d14c", "sd")],
  cbind(sra.2019.inc.df, sd = sd.inc.fx(sra.2019.inc.df))[ , c("PM", "ECO", "PMeco", "lyr_bot", "d14c", "sd")],
  by = c("PM", "ECO", "PMeco", "lyr_bot"),
  suffixes = c("_01", "_19"))
sra.19.01.inc.d.df$dd14c_01_19 <- sra.19.01.inc.d.df$d14c_01 - sra.19.01.inc.d.df$d14c_19
sra.19.01.inc.d.df$sd14c_01_19 <- sqrt(
  sra.19.01.inc.d.df$sd_01^2/2 + sra.19.01.inc.d.df$sd_19^2/2)

## filter outliers
sra.01.19.inc.d.df <- sra.01.19.inc.d.df[-which(sra.01.19.inc.d.df$d14c < -200), ]

# df by depth
sra.01.19.inc.d.10.df <- sra.01.19.inc.d.df[sra.01.19.inc.d.df$lyr_bot == 10, ]
sra.01.19.inc.d.20.df <- sra.01.19.inc.d.df[sra.01.19.inc.d.df$lyr_bot == 20, ]
sra.01.19.inc.d.30.df <- sra.01.19.inc.d.df[sra.01.19.inc.d.df$lyr_bot == 30, ]
```

# Introduction

*	Parent material key factor affecting radiocarbon content of bulk soil
* Development and persistence of poorly crystalline minerals associated with depleted ^14^C content of soil organic matter, i.e. soil C persistence
*	Decomposition rates increase with temperature, suggesting ^14^C of heterotrophically respired CO~2~ will be closer to atmospheric ^14^C in warmer sites than in colder sites 
*	Climate and parent material interact on geologic timescales: weathering leads to formation of poorly crystalline minerals but eventually to degradation of these minerals over time, e.g. poorly crystalline mineral content is highest in intermediate stages of soil development
* How do these interactions affect the balance of C stocks stored in soils?

*[NB: should show C stocks, i.e. context for 14C trends]*

# Hypotheses

1. Parent material will explain more of the variance in the change over time for bulk soil $\Delta$^14^C than climate, because parent material controls $\Delta$^14^C of the slow cycling soil C that dominates the bulk $\Delta$^14^C signal.
2. Climate will explain more of the variance in the change over time for respired soil $\Delta$^14^C-CO~2~ than parent material, as climate controls $\Delta$^14^C of the fast cycling soil C that dominates the respired $\Delta$^14^C-CO~2~ signal.
3. Change in bulk soil $\Delta$^14^C over time will be greatest in soils developed on granite soils and least in soils developed on andesite (GR > BS > AN)
4. Change in respired $\Delta$^14^C-CO~2~ over time will be greatest in the warm climate sites and least in the cold climate sites (warm > cool > cold)
5. Parent material will not affect the change in respired $\Delta$^14^C-CO~2~ over time (AN = BS = GR)

# Methods

I constructed linear models for $\Delta$^14^C of bulk soil C and heterotrophically respired CO~2~ as function of parent material (PM), climate (CL), and year. I considered both the two-way and three-way interactions, using models of the form: lm(d14c ~ PM * CL * year). Prior to modeling I interpolated the 2001 and 2009 $\Delta$^14^C data to the 2019 sampling depths using a mass-preserving spline weighted by carbon stocks. In order to simplify the analysis and focus on the change in $\Delta$^14^C over time as a function of parent material and climate, I modeled each depth increment separately. I focused on the top 30cm only (i.e. 0-10cm, 10-20cm, and 20-30cm depth increments) owing to a lack of data from the 2001 sampling for depths below this depth. 

`r pagebreak()`

# Results
## Depth profiles

>**Fig. `r {(fig.n <- 1)}`. Depth profiles of $\Delta$^14^C of bulk soil and respired CO~2~**

> Bulk soil points show the mean of three replicate profiles. We composited the profile replicates prior to incubation, and respired data points show the mean of laboratory duplicates. Error bars show ±1 SD for bulk soils and the min. and max. for respired CO~2~. Respired CO~2~ from the cold granite soil was extremely depleted in $\Delta$^14^C and thus is excluded for display purposes.

```{r bulk-inc-plot-data, fig.height = 6}
# 2019
load("sra.19.inc.blk.RData")
# 2001
load("sra.01.inc.blk.RData")

## plot
# 2019
# sra.19.inc.blk %>%
#   mutate(ECOtype = paste0(eco, " (", Type, ")")) %>%
#   arrange(lyr_bot) %>%
#   ggplot(., aes(d14c_mean, lyr_bot, color = pm, shape = ECOtype, linetype = Type)) +
#   geom_vline(xintercept = atm.d14.2019) +
#   geom_hline(yintercept = 0) +
#   geom_point(size = 3) +
#   geom_errorbarh(
#     aes(xmin = d14c_l,
#         xmax = d14c_u,
#         color = pm),
#     height = 1.5) +
#   geom_path() +
#   scale_y_reverse() +
#   scale_color_manual(name = "Parent material",
#                      values = c("andesite" = andesite,
#                                 "basalt" = basalt,
#                                 "granite" = granite)) +
#   scale_shape_manual(name = "Ecosystem (type)",
#                      values = c("warm (bulk)" = 15,
#                                 "cool (bulk)" = 16,
#                                 "cold (bulk)" = 17,
#                                 "warm (inc)" = 0,
#                                 "cool (inc)" = 1,
#                                 "cold (inc)" = 2)) +
#   xlab(expression(Delta*''^14*'C (‰)')) +
#   ylab("Depth (cm)") +
#   facet_grid(rows = vars(eco), cols = vars(pm)) +
#   theme_bw() +
#   theme(panel.grid.minor = element_blank())

# 2001
# sra.01.inc.blk %>%
#   mutate(ECOtype = paste0(eco, " (", Type, ")")) %>%
#   ggplot(., aes(d14c_mean, lyr_bot, color = pm, shape = ECOtype, linetype = Type)) +
#   geom_vline(xintercept = atm.d14.2001) +
#   geom_hline(yintercept = 0) +
#   geom_point(size = 3) +
#   geom_errorbarh(
#     aes(xmin = d14c_l, 
#         xmax = d14c_u,
#         color = pm), 
#     height = 1.5) +
#   geom_path() +
#   scale_y_reverse() +
#   scale_color_manual(name = "Parent material",
#                      values = c("andesite" = andesite, 
#                                 "basalt" = basalt, 
#                                 "granite" = granite)) +
#   scale_shape_manual(name = "Ecosystem (type)",
#                      values = c("warm (bulk)" = 15, 
#                                 "cool (bulk)" = 16, 
#                                 "cold (bulk)" = 17,
#                                 "warm (inc)" = 0, 
#                                 "cool (inc)" = 1, 
#                                 "cold (inc)" = 2)) +
#   scale_x_continuous(limits = c(-100, 200)) +
#   xlab(expression(Delta*''^14*'C (‰)')) +
#   ylab("Depth (cm)") +
#   facet_grid(rows = vars(eco), cols = vars(pm)) +
#   theme_bw() +
#   theme(panel.grid.minor = element_blank())

# both
sra.01.inc.blk %>%
  rbind(., sra.19.inc.blk) %>%
  mutate(Type = ifelse(Type == "inc", "respired", "bulk soil")) %>%
  mutate(ECOtype = paste0(eco, " (", Type, ")")) %>%
  filter(lyr_bot > 0) %>%
  ggplot(., aes(d14c_mean, lyr_bot, color = pm, shape = ECOtype, linetype = ECOtype)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  geom_point(size = 3) +
  geom_errorbarh(
    aes(xmin = d14c_l, 
        xmax = d14c_u,
        color = pm), 
    height = 1.5) +
  geom_path() +
  scale_y_reverse() +
  scale_color_manual(name = "Parent material",
                     values = c("andesite" = andesite, 
                                "basalt" = basalt, 
                                "granite" = granite)) +
  scale_shape_manual(name = "Climate (type)",
                     values = c("warm (bulk soil)" = 15, 
                                "cool (bulk soil)" = 16, 
                                "cold (bulk soil)" = 17,
                                "warm (respired)" = 0, 
                                "cool (respired)" = 1, 
                                "cold (respired)" = 2)) +
  scale_linetype_manual(name = "Climate (type)",
                        values = c("warm (bulk soil)" = 1, 
                                   "cool (bulk soil)" = 1, 
                                   "cold (bulk soil)" = 1,
                                   "warm (respired)" = 2, 
                                   "cool (respired)" = 2, 
                                   "cold (respired)" = 2)) +
  coord_cartesian(
    xlim = c(-100, 200), 
    ylim = c(43, 0),
    expand = FALSE) +
  scale_x_continuous(breaks = c(-50, 0, 50, 150)) +
  xlab(expression(Delta*''^14*'C (‰)')) +
  ylab("Depth (cm)") +
  facet_grid(rows = vars(Year), cols = vars(eco)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())

# # plot time series in same panel, with type as layout and splined depths
# # both
# sra.ts.all %>%
#   filter(year != 2009) %>%
#   filter(lyr_bot > 0) %>%
#   mutate(Type = ifelse(Type == "inc", "respired", "bulk soil")) %>%
#   mutate(ecoYear = paste0(eco, " (", year, ")")) %>%
#   ggplot(., aes(d14c, lyr_bot, color = pm, shape = ecoYear, linetype = ecoYear)) +
#   geom_vline(xintercept = 0) +
#   geom_hline(yintercept = 0) +
#   geom_point(size = 3) +
#   geom_errorbarh(
#     aes(xmin = d14c_l,
#         xmax = d14c_u,
#         color = pm),
#     height = .7) +
#   geom_path() +
#   scale_y_reverse() +
#   scale_color_manual(name = "Parent material",
#                      values = c("andesite" = andesite,
#                                 "basalt" = basalt,
#                                 "granite" = granite)) +
#   scale_shape_manual(name = "Ecosystem (type)",
#                      values = c("warm (2001)" = 15,
#                                 "cool (2001)" = 16,
#                                 "cold (2001)" = 17,
#                                 "warm (2019)" = 0,
#                                 "cool (2019)" = 1,
#                                 "cold (2019)" = 2)) +
#   scale_linetype_manual(name = "Ecosystem (type)",
#                         values = c("warm (2001)" = 1,
#                                    "cool (2001)" = 1,
#                                    "cold (2001)" = 1,
#                                    "warm (2019)" = 2,
#                                    "cool (2019)" = 2,
#                                    "cold (2019)" = 2)) +
#   coord_cartesian(
#     xlim = c(-100, 200),
#     ylim = c(32, 8),
#     expand = FALSE) +
#   scale_x_continuous(breaks = c(-50, 0, 50, 150)) +
#   xlab(expression(Delta*''^14*'C (‰)')) +
#   ylab("Depth (cm)") +
#   facet_grid(rows = vars(Type), cols = vars(eco)) +
#   theme_bw() +
#   theme(panel.grid.minor = element_blank())
```

## ANOVA
### Bulk soil

```{r models-blk-sum-con, include = FALSE}
# 0-10
blk.PMeco.10.lm <- lm(d14c ~ PM * ECO * year, sra.01.09.19.d.10df)
summary(blk.PMeco.10.lm)
blk.PMeco.10.et <- emtrends(blk.PMeco.10.lm, pairwise ~ PM * ECO, var = "year")
# ordered trends
# c(-6.1, -5, -2.8, -2.3, -2.1, -1.9, -.3, .1, .8)

# 10-20
blk.PMeco.20.lm <- lm(d14c ~ PM * ECO * year, sra.01.09.19.d.20df)
summary(blk.PMeco.20.lm)
blk.PMeco.20.et <- emtrends(blk.PMeco.20.lm, pairwise ~ PM * ECO, var = "year")
# ordered trends
# c(-3.7, -3.7, -2.1, -1.1, -.3, -.2, .42, .43, 2.1)

# 20-30
blk.PMeco.30.lm <- lm(d14c ~ PM * ECO * year, sra.01.09.19.d.30df)
summary(blk.PMeco.30.lm)
blk.PMeco.30.et <- emtrends(blk.PMeco.30.lm, pairwise ~ PM * ECO, var = "year")
# ordered trends
# c(-6.4, -3.6, -1, .2, .3, .3, .3, 1.1, 1.6)
```

Both parent material and climate are significant factors for explaining the variance in bulk soil $\Delta$^14^C at all depths (Table 1, Table 2, Table 3). Climate explains more of the variance at the surface (0-10cm) than does parent material (`r round(anova(blk.PMeco.10.lm)[["Sum Sq"]][2]/sum(anova(blk.PMeco.30.lm)[["Sum Sq"]])*100,1)`% and `r round(anova(blk.PMeco.10.lm)[["Sum Sq"]][1]/sum(anova(blk.PMeco.30.lm)[["Sum Sq"]])*100,1)`%, for climate and parent material, respectively) (Table 1). Below 10cm in depth the relative importance of these two factors reverses, with parent material explaining more variance than climate (Table 2). At the deepest depth (20-30cm) parent material explains `r round(anova(blk.PMeco.30.lm)[["Sum Sq"]][1]/sum(anova(blk.PMeco.30.lm)[["Sum Sq"]])*100,1)`% of variance in bulk soil $\Delta$^14^C compared to `r round(anova(blk.PMeco.30.lm)[["Sum Sq"]][2]/sum(anova(blk.PMeco.30.lm)[["Sum Sq"]])*100, 1)`% for climate (Table 3). The two-way interaction between parent material and climate is signficant in the top 20cm (*p* < 0.05), but not at the deepest depth.

Year is highly significant at the surface (*p* < 0.0001, Table 1), but is only marginally significant in the deeper soil layers (*p* < 0.10) (Table 2, Table 3). Interestingly, the two-way interaction between climate and year is much more significant than the interaction between parent material and year at all depths. This appears to be driven by the strong decrease in cool site $\Delta$^14^C over time relative to the warm or cold sites (**Fig. 16b**). 

Overall, the models explains less of the variance in the data deeper in the soil: 73%, 61%, and 56% for 0-10cm, 10-20cm, and 20-30cm, respectively (Table 1, Table 2, Table 3).

*NB in subsequent tables "ECO" = climate*

#### 0-10cm
>**Table `r {(tab.n <- 1)}`. ANOVA: Bulk soil, 0-10cm**

```{r models-blk-10-anova}
anova(blk.PMeco.10.lm)
```

#### 10-20cm
>**Table `r {(tab.n <- tab.n + 1)}`. ANOVA: Bulk soil, 10-20cm**

```{r models-blk-20-anova}
anova(blk.PMeco.20.lm)
```

#### 20-30cm
>**Table `r {(tab.n <- tab.n + 1)}`. ANOVA: Bulk soil, 20-30cm**

```{r models-blk-30-anova}
anova(blk.PMeco.30.lm)
```

### Respired CO~2~

The relative importance of parent material and climate for explaining the variance in $\Delta$^14^C-CO~2~ changes with depth in a similar manner as for bulk soil. Climate explains more of the variance in $\Delta$^14^C-CO~2~ at the surface (0-10cm) than does parent material (Table 4). Unlike bulk soil, however, parent material is not significant for $\Delta$^14^C-CO~2~ at the surface (*p* = ). Both parent material and climate are significant (*p* < 0.0005) for $\Delta$^14^C-CO~2~ at the 10-20cm depth, but parent material explains more variance than does climate (Table 5). At the deepest depth (20-30cm) climate is no longer a siginficant predictor of $\Delta$^14^C-CO~2~ at all, while parent material is strongly siginficant (*p* = 0.006) (Table 6). 

Year is highly significant for 0-10cm and 10-20cm (p < 0.0001), but not significant at 20-30cm (p = 0.0950). The two-way interaction between climate and year is significant at all depths, while the two-way interaction between parent material and year is siginficant below 10cm. 

We were able to explain more of the total variance in $\Delta$^14^C-CO~2~ than in bulk soil $\Delta$^14^C. The models explained a similar amount of variance in $\Delta$^14^C-CO~2~ at the 0-10cm and 10-20cm depths (R^2^ of 0.86 and 0.89, respectively), and had a similar amount of unexplained variance at 20-30cm (R^2^ = 0.70) as the model for bulk soil $\Delta$^14^C at the surface (R^2^ = 0.73).

#### 0-10cm
```{r models-inc-10-sum-con, include = FALSE}
# 0-10cm
inc.PMeco.10.lm <- lm(d14c ~ PM * ECO * year, sra.01.19.inc.d.10.df)
summary(inc.PMeco.10.lm)
inc.PMeco.10.et <- emtrends(inc.PMeco.10.lm, pairwise ~ PM * ECO, var = "year")
# ordered trends
# c(-6.2, -4.9, -3.9, -3.7, -3.1, -2.9, -2.3, -1.4, .1)

# 10-20
inc.PMeco.20.lm <- lm(d14c ~ PM * ECO * year, sra.01.19.inc.d.20.df)
summary(inc.PMeco.20.lm)
inc.PMeco.20.et <- emtrends(inc.PMeco.20.lm, pairwise ~ PM * ECO, var = "year")
# ordered trends
# c(-5.9, -4.3, -3.9, -2.2, -1.1, -1, -.8, 1.4, 4.8)

# 20-30
inc.PMeco.30.lm <- lm(d14c ~ PM * ECO * year, sra.01.19.inc.d.30.df)
summary(inc.PMeco.30.lm)
inc.PMeco.30.et <- emtrends(inc.PMeco.30.lm, pairwise ~ PM * ECO, var = "year")
# ordered trends
# c(-7.8, -3.5, -1.5, .1, .5, 1.4, 1.5, 2.8, 9.7)
```
>**Table `r {(tab.n <- tab.n + 1)}`. ANOVA: Respired CO~2~, 0-10cm**

```{r models-inc-10-anova}
anova(inc.PMeco.10.lm)
```

#### 10-20cm
>**Table `r {(tab.n <- tab.n + 1)}`. ANOVA: Respired CO~2~, 10-20cm**

```{r models-inc-20-anova}
anova(inc.PMeco.20.lm)
```

#### 20-30cm
>**Table `r {(tab.n <- tab.n + 1)}`. ANOVA: Respired CO~2~, 20-30cm**

```{r models-inc-30-anova}
anova(inc.PMeco.30.lm)
```

## Marginal mean trends: 2001-2019

The next set of plots show least-square mean estimates of trends within parent material and climate groups for the time period 2001 to 2019. Plots are shown for each depth increment. The first two plots show the full model with interactions between parent material and climate. The same curves are shown twice, but grouped differently to emphasize differences across each level of the grouping factors. Finally, I show the estimated trend and associated confidence intervals for the change in $\Delta$^14^C over time averaged across either parent material or climate. Raw data are shown as points.

Following each plot I show the contrasts for the slope estimates within parent material and climate groups. The first set of numbers show marginal means estimated within groups; second set of numbers show the contrasts, i.e. are the trends over time significantly different between different parent materials or climate conditions?

`r pagebreak()`

### Bulk soil
#### 0-10cm
>**Fig. `r {(fig.n <- fig.n + 1)}`. ^a)^ parent material**

```{r plot-est-slopes-bulk-10-intPM}
# PM | ECO
emmip.plot.fx(lyr_bot = 10,
              Type = "Bulk",
              var = "PM",
              interact_var = "ECO")
```
>**^b)^ ecosystem**

```{r plot-est-slopes-bulk-10-intECO}
# ECO | PM
emmip.plot.fx(lyr_bot = 10,
              Type = "Bulk",
              var = "ECO",
              interact_var = "PM")
```

We can see the strong effect of climate on the temporal trend in bulk soil $\Delta$^14^C clearly in Fig. `r {fig.n}`. ^a)^. Curves are much steeper at the warm site for all three parent materials, while they are nearly flat at the coldest site. However, at the cool site we can also see the effect of parent material within the climate grouping. The steepest slope and the most enriched $\Delta$^14^C is seen in the granite soils, while the andesite soils show no change in $\Delta$^14^C over time and are also much less enriched. Overall, rates of change in granite and basalt soils were not significantly different in any of the ecosystems. However, the andsite soils are changing significantly faster than the basalt soils at the warmest sites and significantly slower than the granite soils at the cool sites (Table 7). 

For the andesite soils we see faster changes over time for the warm site relative to the cool and cold sites when considering the contrasts between climates within parent materials (*p* = 0.0006, and *p* = 0.0657 respectively) (Fig. 5b, Table 8). For the granite soils, however, we see significantly (*p* = 0.0105) faster rates of change at the cool sites than at either the warm or the cold sites. 

```{r plot-est-slopes-bulk-10-PM, include = FALSE}
# PM
emmip.plot.fx(lyr_bot = 10,
              Type = "Bulk",
              var = "PM")
```
```{r plot-est-slopes-bulk-10-ECO, include = FALSE}
# ECO
emmip.plot.fx(lyr_bot = 10,
              Type = "Bulk",
              var = "ECO")
```

`r pagebreak()`
#### Trends & Contrasts
>**Table `r {(tab.n <- tab.n + 1)}`. Bulk soil, 0-10cm: Parent material trends within climate**

```{r trends-est-slopes-bulk-10-PMeco}
print("Interactions, PM | ECO")
emtrends(blk.PMeco.10.lm, pairwise ~ PM | ECO, var = "year")
```
`r pagebreak()`
>**Table `r {(tab.n <- tab.n + 1)}`. Bulk soil, 0-10cm: Climate trends within parent material**

```{r trends-est-slopes-bulk-10-ECOpm}
print("Interactions, ECO | PM")
emtrends(blk.PMeco.10.lm, pairwise ~ ECO | PM, var = "year")
```
`r pagebreak()`

### 10-20cm
#### Interaction plots
>**Fig. `r {(fig.n <- fig.n + 1)}`. ^a)^ parent material** 

```{r plot-est-slopes-bulk-20-intPM}
# PM | ECO
emmip.plot.fx(lyr_bot = 20,
              Type = "Bulk",
              var = "PM",
              interact_var = "ECO")
```
**^b)^ climate**

```{r plot-est-slopes-bulk-20-intECO}
# ECO | PM
emmip.plot.fx(lyr_bot = 20,
              Type = "Bulk",
              var = "ECO",
              interact_var = "PM")
```

text

```{r plot-est-slopes-bulk-20-PM, include = FALSE}
# PM
emmip.plot.fx(lyr_bot = 20,
              Type = "Bulk",
              var = "PM")
```
```{r plot-est-slopes-bulk-20-ECO, include = FALSE}
# ECO
emmip.plot.fx(lyr_bot = 20,
              Type = "Bulk",
              var = "ECO")
```
`r pagebreak()`

#### Trends & Contrasts
>**Table `r {(tab.n <- tab.n + 1)}`.  Bulk soil, 10-20cm: Parent material trends within climate**

```{r trends-est-slopes-bulk-20-PMeco}
print("Interactions, PM | ECO")
emtrends(blk.PMeco.20.lm, pairwise ~ PM | ECO, var = "year")
```
`r pagebreak()`
>**Table `r {(tab.n <- tab.n + 1)}`. Bulk soil, 10-20cm: Climate trends within parent material**

```{r trends-est-slopes-bulk-20-ECOpm}
print("Interactions, ECO | PM")
emtrends(blk.PMeco.20.lm, pairwise ~ ECO | PM, var = "year")
```

`r pagebreak()`
### 20-30cm
#### Interaction plots
>**Fig. `r {(fig.n <- fig.n + 1)}`. ^a)^ parent material**

```{r plot-est-slopes-bulk-30-intPM}
# PM | ECO
emmip.plot.fx(lyr_bot = 30,
              Type = "Bulk",
              var = "PM",
              interact_var = "ECO")
```
**^b)^ climate**

```{r plot-est-slopes-bulk-30-intECO}
# ECO | PM
emmip.plot.fx(lyr_bot = 30,
              Type = "Bulk",
              var = "ECO",
              interact_var = "PM")
```

text

```{r plot-est-slopes-bulk-30-PM, include = FALSE}
# PM
emmip.plot.fx(lyr_bot = 30,
              Type = "Bulk",
              var = "PM")
```
```{r plot-est-slopes-bulk-30-ECO, include = FALSE}
# ECO
emmip.plot.fx(lyr_bot = 30,
              Type = "Bulk",
              var = "ECO")
```
`r pagebreak()`

#### Trends & Contrasts
>**Table `r {(tab.n <- tab.n + 1)}`. Bulk soil, 20-30cm: Parent material trends within climate**

```{r trends-est-slopes-bulk-30-PMeco}
print("Interactions, PM | ECO")
emtrends(blk.PMeco.30.lm, pairwise ~ PM | ECO, var = "year")
```
`r pagebreak()`
>**Table `r {(tab.n <- tab.n + 1)}`. Bulk soil, 20-30cm: Climate trends within parent material**

```{r trends-est-slopes-bulk-30-ECOpm}
print("Interactions, ECO | PM")
emtrends(blk.PMeco.30.lm, pairwise ~ ECO | PM, var = "year")
```
`r pagebreak()`

## Respired CO~2~
### 0-10cm
#### Interaction plots
>**Fig. `r {(fig.n <- fig.n + 1)}`. ^a)^ parent material**

```{r plot-est-slopes-inc-10-intPM}
# PM | ECO
emmip.plot.fx(lyr_bot = 10,
              Type = "Respired",
              var = "PM",
              interact_var = "ECO")
```
**^b)^ climate**

```{r plot-est-slopes-inc-10-intECO}
# ECO | PM
emmip.plot.fx(lyr_bot = 10,
              Type = "Respired",
              var = "ECO",
              interact_var = "PM")
```

text

```{r plot-est-slopes-inc-10-PM, include = FALSE}
# shows data averaged across PM, ECO---not shown as all interactions highly significant
# PM
emmip.plot.fx(lyr_bot = 10,
              Type = "Respired",
              var = "PM")
```
```{r plot-est-slopes-inc-10-ECO, include = FALSE}
# ECO
emmip.plot.fx(lyr_bot = 10,
              Type = "Respired",
              var = "ECO")
```
`r pagebreak()`
#### Trends & Contrasts
>**Table `r {(tab.n <- tab.n + 1)}`. Respired  CO~2~, 0-10cm: Parent material trends within climate**

```{r trends-est-slopes-inc-10}
print("Interactions, PM | ECO")
emtrends(inc.PMeco.10.lm, pairwise ~ PM | ECO, var = "year")
```
`r pagebreak()`
>**Table `r {(tab.n <- tab.n + 1)}`. Respired  CO~2~, 0-10cm: Climate trends within parent material**

```{r}
print("Interactions, ECO | PM")
emtrends(inc.PMeco.10.lm, pairwise ~ ECO | PM, var = "year")
```
`r pagebreak()`

### 10-20cm
#### Interaction plots
>**Fig. `r {(fig.n <- fig.n + 1)}`. ^a)^ parent material**

```{r plot-est-slopes-inc-20-intPM}
# PM | ECO
emmip.plot.fx(lyr_bot = 20,
              Type = "Respired",
              var = "PM",
              interact_var = "ECO")
```
**^b)^ climate**

```{r plot-est-slopes-inc-20-intECO}
# ECO | PM
emmip.plot.fx(lyr_bot = 20,
              Type = "Respired",
              var = "ECO",
              interact_var = "PM")
```

text

```{r plot-est-slopes-inc-20-PM, include = FALSE}
# PM
emmip.plot.fx(lyr_bot = 20,
              Type = "Respired",
              var = "PM")
```
```{r plot-est-slopes-inc-20-ECO, include = FALSE}
# ECO
emmip.plot.fx(lyr_bot = 20,
              Type = "Respired",
              var = "ECO")
```
`r pagebreak()`
#### Trends & Contrasts
>**Table `r {(tab.n <- tab.n + 1)}`. Respired  CO~2~, 10-20cm: Parent material trends within climate**

```{r trends-est-slopes-inc-20-PMeco}
print("Interactions, PM | ECO")
emtrends(inc.PMeco.20.lm, pairwise ~ PM | ECO, var = "year")
```
`r pagebreak()`
>**Table `r {(tab.n <- tab.n + 1)}`. Respired  CO~2~, 10-20cm: Climate trends within parent material**

```{r trends-est-slopes-inc-20-ECOpm}
print("Interactions, ECO | PM")
emtrends(inc.PMeco.20.lm, pairwise ~ ECO | PM, var = "year")
```
`r pagebreak()`

### 20-30cm
#### Interaction plots
>**Fig. `r {(fig.n <- fig.n + 1)}`. ^a)^ parent material**

```{r plot-est-slopes-inc-30-intPM}
# PM | ECO
emmip.plot.fx(lyr_bot = 30,
              Type = "Respired",
              var = "PM",
              interact_var = "ECO")
```
**^b)^ climate**

```{r plot-est-slopes-inc-30-intECO}
# ECO | PM
emmip.plot.fx(lyr_bot = 30,
              Type = "Respired",
              var = "ECO",
              interact_var = "PM")
```

text

```{r plot-est-slopes-inc-30-PM, include = FALSE}
# PM
emmip.plot.fx(lyr_bot = 30,
              Type = "Respired",
              var = "PM")
```
```{r plot-est-slopes-inc-30-ECO, include = FALSE}
# ECO
emmip.plot.fx(lyr_bot = 30,
              Type = "Respired",
              var = "ECO")
```
`r pagebreak()`

#### Trends & Contrasts
>**Table `r {(tab.n <- tab.n + 1)}`. Respired  CO~2~, 20-30cm: Parent material trends within climate**

```{r trends-est-slopes-inc-30-PMeco}
print("Interactions, PM | ECO")
emtrends(inc.PMeco.30.lm, pairwise ~ PM | ECO, var = "year")
```
`r pagebreak()`
>**Table `r {(tab.n <- tab.n + 1)}`. Respired  CO~2~, 20-30cm: Climate trends within parent material**

```{r trends-est-slopes-inc-30-ECOpm}
print("Interactions, ECO | PM")
emtrends(inc.PMeco.30.lm, pairwise ~ ECO | PM, var = "year")
```

`r pagebreak()`
## Interactions with mineral assemblages

Selective dissolution of iron and aluminum bearing secondary minerals provides quantitative information about the differences in mineral assemblages and the nature of mineral-organic associations among different soils (Sposito 2008). 
- Oxalate extractable Fe and Al
  - targets min-OM complexes
  - Fe_ox extracts SRO iron oxyhydroxides, e.g. ferrihydrite
  - Al_ox extracts SRO aluminosilicates, e.g. allophane and imogolite
    - NB: the concentration of the SRO aluminosilicates can be estimated from the difference between Al_ox and Al_py (Parfitt & Childs 1988)
  - relatively tighly bound?
- Dithionate citrate extractable Fe
  - targets iron OM complexes
  - extracts iron from both crystalline (goethite, hematite) and non-crystalline secondary oxyhydroxides (ferrihydrite)
    - non-crytalline component is therefore: Fe_dc - Fe_ox (Parfitt & Childs 1988)
- Pyrophosphate extractable Al
  - targets min-OM complexes formed by cation bridging (LaCroix et al. 2013, Bruckhart 1982, Jansen et al. 2011); of particular note for smectite rich soils, i.e. 2:1 clays (Wattel-Koekkoek et al. 2001)
  - particularly enriched in carboxylic and carbonyl C (Ellerbrock et al 2018)
  - relatively exchangeable pool
  
The differences in $\Delta$^14^C between bulk soil and heterotrophically respired CO~2~ indicates the presence of persistent soil C pools that are inaccessible to the microbial community. Greater differences can indicate the presence of 
  
```{r plot-min-14c}
load("sra.all.min.RData")
sra.all.min %>%
  mutate(pmEcoDepth = paste0(pm, eco, lyr_bot),
         ecoYear = paste0(eco, " (", year, ")"),
         width = ifelse(min == "Al_py" | min == "Fe_ox", .3, 1.5)) %>%
  filter(pmEcoDepth != "granitecold30" & pmEcoDepth != "granitecold20") %>%
  ggplot(., aes(conc, blk.inc)) +
  geom_point(aes(color = pm, shape = ecoYear, size = depth)) +
  geom_errorbar(
    aes(ymin = blk.inc - blk.inc.sd,
        ymax = blk.inc + blk.inc.sd,
        color = pm,
        width = width)) +
  scale_color_manual(name = "Parent material",
                     values = c("andesite" = andesite,
                                "basalt" = basalt,
                                "granite" = granite)) +
  scale_shape_manual(name = "Climate (year)",
                     values = c("warm (2001)" = 15, 
                                "cool (2001)" = 16, 
                                "cold (2001)" = 17,
                                "warm (2019)" = 0, 
                                "cool (2019)" = 1, 
                                "cold (2019)" = 2)) +
  scale_size_manual(name = "Depth",
                    values = (3:5)) +
  facet_wrap(vars(min), scales = "free") +
  ylab(expression('Bulk - Respired '*Delta*''^14*'C (‰)')) +
  xlab(expression('Concentration (g kg'^-1*')')) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())

# Fe_dc alone
#####
# warm
sra.all.min %>%
  filter(eco == "warm" & min == "Fe_dc") %>%
  mutate(ecoYear = paste0(eco, " (", year, ")")) %>%
  ggplot(., aes(conc, blk.inc)) +
  geom_point(aes(color = pm, shape = ecoYear, size = depth)) +
  geom_errorbar(
    aes(ymin = blk.inc - blk.inc.sd,
        ymax = blk.inc + blk.inc.sd,
        color = pm),
        width = 1.5) +
  scale_color_manual(name = "Parent material",
                     values = c("andesite" = andesite,
                                "basalt" = basalt,
                                "granite" = granite)) +
  scale_shape_manual(name = "Climate (year)",
                     values = c("warm (2001)" = 15,
                                "warm (2019)" = 0)) +
  scale_size_manual(name = "Depth",
                    values = (3:5)) +
  facet_wrap(vars(min), scales = "free") +
  ylab(expression('Bulk - Respired '*Delta*''^14*'C (‰)')) +
  xlab(expression('Concentration (g kg'^-1*')')) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())

# cool + cold
sra.all.min %>%
  filter(eco != "warm" & min == "Fe_dc") %>%
  mutate(pmEcoDepth = paste0(pm, eco, lyr_bot),
         ecoYear = paste0(eco, " (", year, ")")) %>%
  filter(pmEcoDepth != "granitecold30" & pmEcoDepth != "granitecold20") %>%
  ggplot(., aes(conc, blk.inc)) +
  geom_point(aes(color = pm, shape = ecoYear, size = depth)) +
  geom_errorbar(
    aes(ymin = blk.inc - blk.inc.sd,
        ymax = blk.inc + blk.inc.sd,
        color = pm),
        width = .5) +
  scale_color_manual(name = "Parent material",
                     values = c("andesite" = andesite,
                                "basalt" = basalt,
                                "granite" = granite)) +
  scale_shape_manual(name = "Climate (year)",
                     values = c("cool (2001)" = 16, 
                                "cold (2001)" = 17,
                                "cool (2019)" = 1, 
                                "cold (2019)" = 2)) +
  scale_size_manual(name = "Depth",
                    values = (3:5)) +
  facet_wrap(vars(min), scales = "free") +
  ylab(expression('Bulk - Respired '*Delta*''^14*'C (‰)')) +
  xlab(expression('Concentration (g kg'^-1*')')) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

```{r min-stats, include = FALSE}
# reshape sra.all.min wider
sra.all.min2 <- sra.all.min %>%
  pivot_wider(names_from = min, values_from = conc)
corrplot(cor(sra.all.min2[sra.all.min2$year == 2001, c(1,4:9)]), method = "number")

## change over time vs. min
# bulk
sra.19.01.d.min <- left_join(
  sra.19.01.d[ , c("lyr_bot", "PM", "ECO", "dd14c_01_19", "sd14c_01_19")], 
  ras18.sp.df[ , c("Al_ox", "Al_py", "Fe_dc", "Fe_ox", "lyr_bot", "PM", "ECO")], 
  by = c("lyr_bot", "PM", "ECO")) 
sra.19.01.d.min2 <- sra.19.01.d.min %>%
  pivot_longer(cols = c("Al_py", "Al_ox", "Fe_ox", "Fe_dc"), names_to = "min", values_to = "conc")

# corr
corrplot(cor(sra.19.01.d.min[ , c(1,4:9)]), method = "number")

# min vs. d/t
sra.19.01.d.min2 %>%
  filter(lyr_bot == 20) %>%
  mutate(eco = factor(ifelse(ECO == "pp", "warm",
                      ifelse(ECO == "wf", "cool", "cold")),
                      levels = c("warm", "cool", "cold")),
         pm = ifelse(PM == "AN", "andesite",
                     ifelse(PM == "BS", "basalt", "granite")),
         width = ifelse(min == "Al_py" | min == "Fe_ox", .3, 1.5)) %>%
  ggplot(., aes(conc, dd14c_01_19)) +
  geom_point(aes(color = pm, shape = eco), size = 3) +
  geom_errorbar(
    aes(ymin = dd14c_01_19 - sd14c_01_19,
        ymax = dd14c_01_19 + sd14c_01_19,
        color = pm,
        width = width)) +
  scale_color_manual(name = "Parent material",
                     values = c("andesite" = andesite,
                                "basalt" = basalt,
                                "granite" = granite)) +
  scale_shape_manual(name = "Climate",
                     values = c("warm" = 15, 
                                "cool" = 16, 
                                "cold" = 17)) +
  facet_wrap(vars(min), scales = "free") +
  ylab(expression('Change in bulk soil '*Delta*''^14*'C (‰) 2019-2001')) +
  xlab(expression('Concentration (g kg'^-1*')')) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())

# inc
sra.19.01.inc.d.min.df <- left_join(
  sra.19.01.inc.d.df[ , c("lyr_bot", "PM", "ECO", "dd14c_01_19", "sd14c_01_19")], 
  ras18.sp.df[ , c("Al_ox", "Al_py", "Fe_dc", "Fe_ox", "lyr_bot", "PM", "ECO")], 
  by = c("lyr_bot", "PM", "ECO"))
sra.19.01.inc.d.min.df2 <- sra.19.01.inc.d.min.df %>%
  pivot_longer(cols = c("Al_py", "Al_ox", "Fe_ox", "Fe_dc"), names_to = "min", values_to = "conc")

# corr
corrplot(cor(sra.19.01.inc.d.min.df[ , c(1,4:9)]), method = "number")

# min vs. d/t
sra.19.01.inc.d.min.df %>%
  filter(lyr_bot == 30) %>%
  mutate(pmeco = paste0(PM, ECO),
         eco = factor(ifelse(ECO == "pp", "warm",
                      ifelse(ECO == "wf", "cool", "cold")),
                      levels = c("warm", "cool", "cold")),
         pm = ifelse(PM == "AN", "andesite",
                     ifelse(PM == "BS", "basalt", "granite")),
         width = ifelse(min == "Al_py" | min == "Fe_ox", .3, 1.5)) %>%
  filter(pmeco != "GRrf") %>%
  ggplot(., aes(conc, dd14c_01_19)) +
  geom_point(aes(color = pm, shape = eco), size = 3) +
  geom_errorbar(
    aes(ymin = dd14c_01_19 - sd14c_01_19,
        ymax = dd14c_01_19 + sd14c_01_19,
        color = pm,
        width = width)) +
  scale_color_manual(name = "Parent material",
                     values = c("andesite" = andesite,
                                "basalt" = basalt,
                                "granite" = granite)) +
  scale_shape_manual(name = "Climate",
                     values = c("warm" = 15, 
                                "cool" = 16, 
                                "cold" = 17)) +
  facet_wrap(vars(min), scales = "free") +
  ylab(expression('Change in respired '*Delta*''^14*'C (‰) 2019-2001')) +
  xlab(expression('Concentration (g kg'^-1*')')) +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```